### 2025-06-09 table1.qmd

---
title: "Table 1"
subtitle: "From Roe et al., 'XRONOS: An Open Data Infrastructure for Archaeological Chronology'"
bibliography: 
  - references.bib
  - data/c14_datasets.bib
execute:
  echo: false
---

This table was omitted from the main manuscript due to formatting restrictions.

```{r deps}
library("dplyr", warn.conflicts = FALSE)
library("glue")
library("gt")
library("readr")
library("stringr")
```

```{r data-c14-datasets}
c14_datasets <- read_tsv("data/c14_datasets.tsv", show_col_types = FALSE)

n_c14_datasets <- nrow(c14_datasets)
n_c14_datasets_sans_publication_year <- nrow(filter(c14_datasets, is.na(publication_year)))
```

```{r tbl-c14-datasets}
#| tbl-cap: Summary of published compilations of radiocarbon dates. For full data, see supplementary materials.
# References in tables broken by:
# https://github.com/quarto-dev/quarto-cli/issues/9342
c14_datasets |>
  transmute(
    database = glue("[{name}]({url})"),
    publication_year, n_dates, citations
  ) |>
  arrange(publication_year) |>
  gt() |>
  cols_label(
    publication_year = "published",
    citations = "references",
    n_dates = "dates"
    ) |>
  cols_label_with(everything(), str_to_sentence) |>
  sub_missing(everything()) |>
  fmt_markdown(c(database, citations)) |>
  fmt_number(c(n_dates), decimals = 0) |>
  cols_width(
    database ~ pct(40),
    publication_year ~ pct(15),
    n_dates ~ pct(15),
    citations ~ pct(30)
  ) |>
  cols_align(
    "left", c(database)
  ) |>
  tab_options(
    table.font.size = 13
  )
```